Sequence Specific Modeling of \emph{E.~coli} Cell-Free Protein Synthesis

### 2017-04-22 Paper_v1.tex
Working on changes in the results

\subsubsection{Energy efficiency}
%The energy efficiency of protein synthesis was linearly related to protein carbon number, both with and without amino acid supplementation (Fig.~\ref{fig:Prof}C-D).
%However, the dependence of energy efficiency on carbon number was more pronounced in the absence of amino acid supplementation.
The optimal energy efficiency of protein synthesis was independent of protein length, with and without amino acid supplementation (Fig.~\ref{fig:Prof}C-D).
The optimal energy efficiency was approximately 86\% for the model proteins sampled.
The optimal energy efficiency was independent of the protein length; $m_{Y}\times(C_{POI})+b_{Y}$ where $m_{Y} =  -4.01\times10^{-4}$ energy efficiency (\%)/carbon number with supplementation,
and $m_{Y} = 3.03\times10^{-3}$ energy efficiency (\%)/carbon number without supplementation.
The y-intercept $b_{Y} = 86.20$ energy efficiency (\%) with supplementation, and $b_{Y} = 67.40$ energy efficiency (\%) without supplementation.
In the presence of amino acids, energy was utilized to power CFPS instead of synthesizing amino acids, thus a constant level of energy efficiency was observed regardless of the protein size.
In the absence of supplementation, there was a drop in energy efficiency, ranging from 68\% to 76\%.
In this case, glucose consumption more than doubled (64\% increase for CAT) compared to cases supplemented with amino acids, meanwhile the productivity was similar for each protein (Fig.~\ref{fig:Prof}D).
Therefore, the energy burden required for synthesizing each amino acid, and powering CFPS, lowered the energy efficiency.
Surprisingly, without amino acid supplementation, proteins with a higher carbon number had a marginally higher energy efficiency;
however, this linear trend was mostly independent of protein size (R\textsuperscript{2} = 0.65).
MBP showed good agreement with the linear efficiency model with and without amino acid supplementation.
The estimated MBP energy efficiency had a maximum error of 5\% without supplementation, and an error of 1\% - 3\% in the presence of amino acid.
The experimentally constrained case of CAT production showed even a lower energy efficiency of 13.3 $\pm$ 4.7\% compared to the theoretical maximum of 84.2 $\pm$ 0.1\%.
This suggested CFPS was not energy optimized and could be improved: first, the experimental measurements showed accumulation of certain amino acids; second,
there was significant accumulation of organic acids, suggesting inefficient energy utilization.

@@

\subsection{Global sensitivity analysis}
To better understand the effect of substrate utilization and the transcription/translation parameters on CFPS performance, we performed global sensitivity analysis on the productivity, energy efficiency and carbon yield for the CAT protein (Fig.~\ref{fig:SI}).
Surprisingly, RNAP and ribosome abundance had only a modest effect;
the translation elongation rate had the largest effect on protein productivity.
On the other hand, oxygen followed by substrate consumption had the largest influence on the energy efficiency and carbon yield, respectively.
The significance of transcription/translation parameters was robust to amino acid supplementation, with the translation rate being the most sensitive across all cases (Fig.~\ref{fig:SI}A).
This suggested that the translation elongation rate, and not transcriptional parameters, controlled productivity.
Underwood and coworkers showed that an increase in ribosome levels did not significantly increase protein yields or rates; however, adding elongation factors increased protein synthesis rates by 33\% and yields by 23\% at 30 minutes \cite{2005_underwood_biotech}.
In addition, Li et al. increased the productivity of firefly luciferase production in PURE CFPS by 5-fold by adjusting factors that affect transcription and translation such as elongation factors, ribosome recycling factor, release factors, chaperones, BSA, and tRNAs \cite{2014_li_PlosOne}.
In examining substrate utilization, glucose consumption was not important for productivity in the presence of amino acid supplementation.
However, its importance increased significantly when amino acids were not available.
On the other hand, amino acid consumption was only sensitive when amino acids synthesis reactions were blocked, as it was the only source of amino acids for CAT synthesis.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-3-Sensitivity-Analysis.pdf}
\caption{Total order sensitivity of deGFP productivity (A) and energy efficiency (B) to specific uptake rates and transcription/translation parameters for three cases: amino acid uptake and synthesis (black), amino acid uptake without synthesis (dark grey), and amino acid synthesis without uptake (light gray). Error bars represent a 95\% CI on the sensitivity index.}
\label{fig:SI}
\end{figure}

@@

\subsection{Optimal Metabolic Flux Distribution}
Amino acid supplementation altered the optimal metabolic flux distribution predicted for CAT production (Fig.~\ref{fig:flux}).
To investigate the influence of amino acid supplementation, we compared the simulated metabolic flux distributions for CAT production with and without external amino acids.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-5-FluxDistribition-Optimal.pdf}
\caption{Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and oxidative phosphorylation, for three different cases: (A) amino acid uptake and synthesis, (B) amino acid uptake without synthesis, and (C) amino acid synthesis without uptake. Mean flux across the ensemble, normalized to glucose uptake flux. Thick arrows indicate flux to or from amino acids.}
\label{fig:flux}
\end{figure}
In the presence of amino acid supplementation, and \textit{de~novo} amino acid synthesis, there was an incomplete TCA cycle, where a combination
of glucose and amino acids powered protein expression (Fig.~\ref{fig:flux}A).
Glucose was consumed to produce acetyl-coenzyme A, and associated by-products,
while glutamate was converted to alpha-ketoglutarate which traveled to oxaloacetic acid and pyruvate for additional amino acid biosynthesis.
On the other hand, in the presence of amino acid supplementation, but without \textit{de~novo} amino acid biosynthesis, there was no TCA cycle flux.
In this case, ATP was produced by a combination of substrate level and oxidative phosphorylation, where
ubiquinone was regenerated via \textit{nuo} activity, without relying on succinate dehydrogenase in the TCA cycle (Fig.~\ref{fig:flux}B).
These first two cases where amino acids were available had similar performance, and their respective metabolic flux distributions had a 99\% correlation for all proteins.
In the absence of amino acid supplementation (where all amino acids were synthesized \textit{de~novo} from glucose),
the energy efficiency and carbon yield decreased; in this case the TCA cycle was largely complete and there was diversion of metabolic flux into the Entner-Doudoroff pathway to produce NADPH (Fig.~\ref{fig:flux}C).
However, each of these three cases represent the theoretical optimum under different conditions; in order to more accurately describe system performance, we constrained the feasible solution space with experimental measurements.

@@

The case constrained by experimental measurements (Fig. \ref{fig:flux_exp}) had the highest correlation of 0.66 with the flux distribution from the case supplied with no amino acids and a correlation of 0.52 with the cases supplied with amino acids.
Thus there are some differences in the flux distribution compared to the optimum solutions which may provide some insight to improve CFPS performance.
Metabolic fluxes were constrained by experimental measurements (available in Supporting Information) where available for the first hour
which constrained the feasible solution space.
The central carbon organic acids showed good agreement with the data with a coefficient of determination of R\textsuperscript{2}=0.92 (Fig.~\ref{fig:flux_exp}B).
Only certain amino acid synthesis reactions were blocked, since during the growth of \emph{E.~coli} not all amino acids were supplied (see Materials and Methods).
During the cell-free reaction all amino acids were supplied, however glucose still traveled through all the major pathways, and the same metabolic precursors were still utilized for amino acid biosynthesis.
In this case, it is unclear which substrate (glucose or amino acids) is used to power CFPS and may in fact be a combination of both.
The optimum solutions only produced the required amount of amino acids necessary, however in examining the measurements there is an accumulation of alanine and glutamine which may explain some of the differences in the correlations of the flux distributions.
Accumulation of pyruvate, lactate, acetate, and other organic acids can be seen (Fig.~\ref{fig:flux_exp}B) and is an inefficiency of carbon utilization.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./figs/Fig-6-FluxDistribition-Experimental.pdf}
\caption{Constraint based simulation of CAT production for an experimentally constrained case. (A) Flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, and oxidative phosphorylation.  Mean flux across the ensemble, normalized to glucose uptake flux. Thick arrows indicate flux to or from amino acids. (B) Central carbon metabolite measurements versus simulations over a one hour time course.}
\label{fig:flux_exp}
\end{figure}

\subsubsection{Potential Alternative Metabolic Optima}
Alternative optimal solutions have the same objective value, e.g., productivity, but different metabolic flux distributions.
Several techniques, e.g., flux variability analysis (FVA) \cite{Mahadevan2003264,Schuetz119} or mixed-integer approaches \cite{LEE2000711} can estimate alternative optima.
In this study, we used flux variability analysis to estimate potential alternative optimal solutions for CAT production constrained by experimental measurements (Fig.~\ref{fig:norm}).
FVA identified pathways required for CAT production; for example, deletion of the glycolysis/gluconoeogenesis or oxidative phosphorylation pathways resulted in no CAT production.
Likewise, there were pathway knockouts that had no effect on productivity or the metabolic flux distribution, such as removal of isoleucine, leucine, histidine and valine biosynthesis.
Globally, the constraint based simulation reached the same optimal CAT productivity for 40\% of the pairwise knockouts, but had different flux distributions compared with the absence of knockouts.
For example, one of the features of the predicted optimal metabolic flux distribution was a high flux through the Entner-Douodoroff (ED) pathway.
Removal of the ED pathway had no effect on the CAT productivity compared to the absence of knockouts (Fig.~\ref{fig:norm}A).
Pairwise knockouts of the ED pathway and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) also resulted in the same optimal CAT productivity.
However, there was a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); thus, alternative optimal metabolic flux distributions exist for CAT production, despite experimental constraints.
In addition, knockouts of amino acid biosynthesis reactions had no effect on the productivity with the exception of ALA, ASP, ASN and GLU, GLN biosynthesis reactions, since amino acids were supplemented in the media.
ssFBA rerouted the metabolic flux to have the same productivity but an alterantive flux distribution (alternative optima), as can be seen for a pairwise knockout of GLY, SER biosynthesis with HIS biosynthesis (Fig.~\ref{fig:norm}B).
Ultimately, to determine the metabolic flux distribution occurring in CFPS, we need to add additional constraints to the flux estimation calculation.
For example, thermodynamic feasibility constraints may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013},
and $^{13}$C labeling in CFPS could provide significant insight.
However, while $^{13}$C labeling techniques are well established for \emph{in~vivo} processes \cite{Zamboni:2009},
application of these techniques to CFPS remains an active area of research.
\begin{figure}[t!]
\includegraphics[width=1.0\textwidth]{./figs/Fig-7-FluxDistribution-Analysis.pdf}
\caption{The norm effect of pairwise knockouts of subgroups in the cell-free network (A) CAT productivity norm compared to no knockouts in the experimentally constrained case. (B) Flux distribution norm compared to no knockouts in the experimentally constrained case.}
\label{fig:norm}
\end{figure}

@@

\subsection*{Formulation and solution of the model equations.}
The sequence-specific flux balance analysis problem was formulated as a linear program:
\begin{equation}
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{X} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\mathcal{L}_{i} \leq w_i \leq \mathcal{U}_{i}  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective cost vector
and $\mathcal{L}_{i}$ and $\mathcal{U}_{i}$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The transcription (T) and translation (X) stoichiometry was modeled based upon the template reactions of Allen and Palsson \cite{Allen:2003aa}:
\begin{table}[!]
\centering
    \caption{Transcription and translation template reactions for protein production.}
    \renewcommand{\arraystretch}{1.2}
    \resizebox{\textwidth}{!}{\begin{tabular}{llcl} \toprule
        \textbf{Description} & \multicolumn{3}{l}{\textbf{Template reaction}} \\ \toprule
        Transcription initiation & $G_\mathcal{P}+R_{T}$ & $\longrightarrow$& $G_{\mathcal{P}}^{*}$ \\
        Transcription ($w_{T}$) & $G_{\mathcal{P}}^{*}+\quad\displaystyle\sum_{\mathclap{k\in\left\{A,C,G,U\right\}}}\eta_{k}\cdot \left(\left\{k\right\}TP + H_{2}O\right)$ & $\longrightarrow$& $mRNA+G_{\mathcal{P}}+R_{T}+\displaystyle\sum_{\mathclap{k\in\left\{A,C,G,U\right\}}} \eta_{k} \cdot PPi$ \\
        mRNA degradation &  $mRNA$ &$\longrightarrow$& \quad $\displaystyle\sum_{\mathclap{k\in\left\{A,C,G,U\right\}}}\eta_{k}\cdot \left\{k\right\}MP$ \\
	Translation initiation & $mRNA+R_{X}$ &$\longrightarrow$& $R_{X}^{*}$ \\
        tRNA charging & $\alpha_{j}\cdot \Big(AA_{j}+tRNA+ATP+H_{2}O\Big)$ &$\longrightarrow$& $\alpha_{j}\cdot \Big(AA_{j}\text{-}tRNA_{j} +AMP+PPi\Big)$ \\
 	&&& $\qquad \qquad \qquad \qquad \quad  {j=1,2,\hdots,20}$\\
        Translation ($w_{X}$) &  $R_{X}^{*}+\displaystyle\sum_{\mathclap{j\in\left\{AA\right\}}}\alpha_{j}\cdot \Big(AA_{j}\text{-} tRNA_{j}+2GTP+2H_{2}O\Big)$ &$\longrightarrow$& $\mathcal{P}+R_{X}+mRNA$\\
&&& $\quad +\displaystyle\sum_{\mathclap{j\in\left\{AA\right\}}}\alpha_{j}\cdot\Big(tRNA+2GDP+2Pi\Big)$ \\ \bottomrule
    \end{tabular}}
\label{tbl:parameters}
\end{table}
where $G_{\mathcal{P}}$ denotes the gene encoding protein product $\mathcal{P}$,
$R_{T}$ denotes the concentration of RNA polymerase,
$G_{\mathcal{P}}^{*}$ denotes the gene bounded by the RNA polymerase (open complex),
$\eta_{i}$ and $ \alpha_{j}$ denote the stoichiometric coefficients for nucleotide and amino acid, respectively,
$Pi$ denotes inorganic phosphate,
$R_{X}$ denotes the ribosome concentration,
$R_{X}^{*}$ denotes bound ribosome,
and $AA_{j}$ denotes $j^{th}$ amino acid.

The objective of the sequence specific flux balance calculation was to maximize the rate of protein translation, $w_{X}$.
The total glucose uptake rate was bounded by [0,40 mM/h] according to experimental data, while the amino acid uptake rates were bounded by [0,30 mM/h], but did not reach the maximum flux.
Gene and protein sequences were taken from literature and are available in the Supporting Information.
The sequence specific flux balance linear program was solved using the GNU Linear Programming Kit (GLPK) v4.55 \cite{GLPK}.
For all cases, amino acid degradation reactions were blocked as these enzymes were knocked out during the cell-free extract preparation \cite{2005_calhoun_BiotechnologyProgress, Garamella:2016aa}.
In the second case, all amino acid synthesis reactions were set to 0 mM/hr since \textit{E. coli} was grown in the presence of amino acids, thus these enzymes would not be present in the cell-free extract media.
In the third case, amino acid uptake reactions were set to 0 mM/hr.
In the experimental constrained case, \textit{E. coli} was grown in the presence of 13 amino acids (alanine, arginine, cysteine, serine, aspartate, glutamate, and glutamine were excluded) \cite{Zawada:2003}, thus the synthesis reactions responsible for those 13 amino acid were set to 0 mM/hr.

@@

\begin{table}[!]
\centering
    \caption{Parameters for sequence specific flux balance analysis}
    \renewcommand{\arraystretch}{1}
    \begin{tabular}{lcccc} \toprule
        \textbf{Description} & \textbf{Parameter} & \textbf{Value} & \textbf{Units} & \textbf{Reference} \\ \toprule
        RNA polymerase concentration & $R_{T}$ & 75 & nM & \cite{Garamella:2016aa} \\
        Ribosome concentration & $R_{X}$ & 1.6 & $\mu$M & \cite{Garamella:2016aa, 2005_underwood_biotech} \\

        & & & & \\
        T7 promoter & $K_{T7}$ & 10 & constant & estimated \\
        Weight RNA polymerase binding alone P70a & $K_{1}$ & 0.014 & constant & estimated \\
        Weight bound RNAP-$\sigma_{70}$ P70a & $K_{2}$ & 10 & constant & estimated \\
        $\sigma_{70}$ concentration & $\sigma_{70}$ & 35 & nM & \cite{Garamella:2016aa} \\
        $\sigma_{70}$ dissociation constant & $K_{D}$ & 130 & nM & \cite{Mauri2014} \\
	      $\sigma_{70}$ hill coefficient & n & 1 & constant & \cite{Mauri2014} \\
        Gene concentration & $G_{\mathcal{P}}$ & 5 & nM &  \cite{Garamella:2016aa} \\ \bottomrule

\subsection*{Calculation of energy efficiency.}
Energy efficiency ($\mathcal{E}$) was calculated as the ratio of protein production to glucose consumption, written in terms of equivalent ATP molecules:
%\begin{equation}\label{eqn:energy-efficiency-definition}
%	\mathcal{E}\equiv\Bigl(\frac{\texttt{q}_{POI}}{\texttt{q}_{GLC}\cdot\texttt{ATP}_{GLC}}\Bigr)\cdot \Bigl(2\times\left(\rm {ATP}_{T}+\rm {CTP}_{T}+\rm {GTP}_{T}+\rm {UTP}_{T}\right)+\rm2\cdot\rm {ATP}_{X}+\rm {GTP}_{X}\Bigl)
%\end{equation}

\begin{equation}\label{eqn:energy-efficiency-definition}
	\mathcal{E}\equiv\frac{w_{X}\left(2\cdot\rm {ATP}_{X}+\rm {GTP}_{X}\right)+w_{T}\left(2\times\left(\rm {ATP}_{T}+\rm {CTP}_{T}+\rm {GTP}_{T}+\rm {UTP}_{T}\right) \right)}{\texttt{q}_{GLC}\cdot\texttt{ATP}_{GLC}}
\end{equation}

where $\rm {ATP}_{T}$, $\rm {CTP}_{T}$, $\rm {GTP}_{T}$, $\rm {UTP}_{T}$ denote the stoichiometric coefficients of each energy species for the transcription of the protein of interest, $\rm {ATP}_{X}$, $\rm {GTP}_{X}$ denote the stoichiometric coefficients of ATP and GTP for the translation of the protein of interest, $\texttt{q}_{GLC}=w_{GLC}$ denotes the glucose uptake rate,
and $\texttt{ATP}_{GLC}$ denotes the equivalent ATP number for glucose produced.
The energy species stoichiometric coefficients are available in the Supporting Information.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{References_v1}

>>>>

\subsubsection{Energy efficiency}
%The energy efficiency of protein synthesis was linearly related to protein carbon number, both with and without amino acid supplementation (Fig.~\ref{fig:Prof}C-D).
%However, the dependence of energy efficiency on carbon number was more pronounced in the absence of amino acid supplementation.
The optimal energy efficiency of protein synthesis was independent of protein length, with and without amino acid supplementation (Fig.~\ref{fig:Prof}C-D).
The optimal energy efficiency was approximately 86\% for the model proteins sampled.
The optimal energy efficiency was independent of the protein length; $m_{Y}\times(C_{POI})+b_{Y}$ where $m_{Y} =  -4.01\times10^{-4}$ energy efficiency (\%)/carbon number with supplementation,
and $m_{Y} = 3.03\times10^{-3}$ energy efficiency (\%)/carbon number without supplementation.
The y-intercept $b_{Y} = 86.20$ energy efficiency (\%) with supplementation, and $b_{Y} = 67.40$ energy efficiency (\%) without supplementation.
In the presence of amino acids, energy was utilized to power CFPS instead of synthesizing amino acids, thus a constant level of energy efficiency was observed regardless of the protein size.
In the absence of supplementation, there was a drop in energy efficiency, ranging from 68\% to 76\%.
In this case, glucose consumption more than doubled (64\% increase for CAT) compared to cases supplemented with amino acids, meanwhile the productivity was similar for each protein (Fig.~\ref{fig:Prof}D).
Therefore, the energy burden required for synthesizing each amino acid, and powering CFPS, lowered the energy efficiency.
Surprisingly, without amino acid supplementation, proteins with a higher carbon number had a marginally higher energy efficiency;
however, this linear trend was mostly independent of protein size (R\textsuperscript{2} = 0.65).
MBP showed good agreement with the linear efficiency model with and without amino acid supplementation.
The estimated MBP energy efficiency had a maximum error of 5\% without supplementation, and an error of 1\% - 3\% in the presence of amino acid.

Experimentally constrained CAT simulations showed suboptimal energy efficiency (Fig.~\ref{fig:Prof}D, dagger).
CAT production was simulated using the constraint based model in combination with experimental measurements of glucose consumption, organic and amino acid consumption and production rates (Fig. ~\ref{fig:network}B).
The experimentally constrained energy efficiency was 13.3 $\pm$ 5.0\% compared to the theoretical maximum of approximately 84 $\pm$ 0.1\%, however, the
CAT productivity was similar between the simulated and measured systems.
This suggested CFPS was not energy optimized.
There were two factors responsible for the difference between the optimal and experimental system,
the glucose consumption rate and the ATP yield per glucose.
The glucose consumption rate was approximately 30 mM/h in the experimental system (even in the presence of amino acids).
However, the constraint based simulation suggested the optimal glucose consumption rate was significantly less than the observed rate, approximately 2 - 4 mM/h.
In the constraint based simulation, the CFPS reaction produced only acetate as a by-product, but in the experimental system acetate, lactate, pyruvate, succinate and malate accumulated during the first hour of production.
Thus, the constraint based simulation was more carbon efficient.
Next, the energy produced per unit glucose was different between the optimal and experimental constrained system.
In the optimal simulation, 12 ATPs were produced per glucose (the theoretical maximum for this network was 21),
while the experimentally constrained simulation produced only 4 ATPs per glucose.
[TODO JV: Think more about this].

@@

\subsection{Global sensitivity analysis}
To better understand the effect of substrate utilization and the transcription/translation parameters on CFPS performance, we performed global sensitivity analysis on the productivity, energy efficiency and carbon yield for the CAT protein (Fig.~\ref{fig:SI}).
Surprisingly, RNAP and ribosome abundance had only a modest effect;
the translation elongation rate had the largest effect on protein productivity.
On the other hand, oxygen followed by substrate consumption had the largest influence on the energy efficiency and carbon yield, respectively.
The significance of transcription/translation parameters was robust to amino acid supplementation, with the translation rate being the most sensitive across all cases (Fig.~\ref{fig:SI}A).
This suggested that the translation elongation rate, and not transcriptional parameters, controlled productivity.
Underwood and coworkers showed that an increase in ribosome levels did not significantly increase protein yields or rates; however, adding elongation factors increased protein synthesis rates by 33\% and yields by 23\% at 30 minutes \cite{2005_underwood_biotech}.
In addition, Li et al. increased the productivity of firefly luciferase production in PURE CFPS by 5-fold by adjusting factors that affect transcription and translation such as elongation factors, ribosome recycling factor, release factors, chaperones, BSA, and tRNAs \cite{2014_li_PlosOne}.
In examining substrate utilization, glucose consumption was not important for productivity in the presence of amino acid supplementation.
However, its importance increased significantly when amino acids were not available.
On the other hand, amino acid consumption was only sensitive when amino acids synthesis reactions were blocked, as it was the only source of amino acids for CAT synthesis.

@@

\subsection{Optimal Metabolic Flux Distribution}
Amino acid supplementation altered the optimal metabolic flux distribution predicted for CAT production (Fig.~\ref{fig:flux}).
To investigate the influence of amino acid supplementation, we compared the simulated metabolic flux distributions for CAT production with and without external amino acids.
In the presence of amino acid supplementation, and \textit{de~novo} amino acid synthesis, there was an incomplete TCA cycle, where a combination
of glucose and amino acids powered protein expression (Fig.~\ref{fig:flux}A).
Glucose was consumed to produce acetyl-coenzyme A, and associated by-products,
while glutamate was converted to alpha-ketoglutarate which traveled to oxaloacetic acid and pyruvate for additional amino acid biosynthesis.
On the other hand, in the presence of amino acid supplementation, but without \textit{de~novo} amino acid biosynthesis, there was no TCA cycle flux.
In this case, ATP was produced by a combination of substrate level and oxidative phosphorylation, where
ubiquinone was regenerated via \textit{nuo} activity, without relying on succinate dehydrogenase in the TCA cycle (Fig.~\ref{fig:flux}B).
These first two cases where amino acids were available had similar performance, and their respective metabolic flux distributions had a 99\% correlation for all proteins.
In the absence of amino acid supplementation (where all amino acids were synthesized \textit{de~novo} from glucose),
the energy efficiency and carbon yield decreased; in this case the TCA cycle was largely complete and there was diversion of metabolic flux into the Entner-Doudoroff pathway to produce NADPH (Fig.~\ref{fig:flux}C).
However, each of these three cases represent the theoretical optimum under different conditions; in order to more accurately describe system performance, we constrained the feasible solution space with experimental measurements.

@@

The case constrained by experimental measurements (Fig. \ref{fig:flux_exp}) had the highest correlation of 0.66 with the flux distribution from the case supplied with no amino acids and a correlation of 0.52 with the cases supplied with amino acids.
Thus there are some differences in the flux distribution compared to the optimum solutions which may provide some insight to improve CFPS performance.
Metabolic fluxes were constrained by experimental measurements (available in Supporting Information) where available for the first hour
which constrained the feasible solution space.
The central carbon organic acids showed good agreement with the data with a coefficient of determination of R\textsuperscript{2}=0.92 (Fig.~\ref{fig:flux_exp}B).
Only certain amino acid synthesis reactions were blocked, since during the growth of \emph{E.~coli} not all amino acids were supplied (see Materials and Methods).
During the cell-free reaction all amino acids were supplied, however glucose still traveled through all the major pathways, and the same metabolic precursors were still utilized for amino acid biosynthesis.
In this case, it is unclear which substrate (glucose or amino acids) is used to power CFPS and may in fact be a combination of both.
The optimum solutions only produced the required amount of amino acids necessary, however in examining the measurements there is an accumulation of alanine and glutamine which may explain some of the differences in the correlations of the flux distributions.
Accumulation of pyruvate, lactate, acetate, and other organic acids can be seen (Fig.~\ref{fig:flux_exp}B) and is an inefficiency of carbon utilization.

\subsubsection{Potential Alternative Metabolic Optima}
Alternative optimal solutions have the same objective value, e.g., productivity, but different metabolic flux distributions.
Several techniques, e.g., flux variability analysis (FVA) \cite{Mahadevan2003264,Schuetz119} or mixed-integer approaches \cite{LEE2000711} can estimate alternative optima.
In this study, we used flux variability analysis to estimate potential alternative optimal solutions for CAT production constrained by experimental measurements (Fig.~\ref{fig:norm}).
FVA identified pathways required for CAT production; for example, deletion of the glycolysis/gluconoeogenesis or oxidative phosphorylation pathways resulted in no CAT production.
Likewise, there were pathway knockouts that had no effect on productivity or the metabolic flux distribution, such as removal of isoleucine, leucine, histidine and valine biosynthesis.
Globally, the constraint based simulation reached the same optimal CAT productivity for 40\% of the pairwise knockouts, but had different flux distributions compared with the absence of knockouts.
For example, one of the features of the predicted optimal metabolic flux distribution was a high flux through the Entner-Douodoroff (ED) pathway.
Removal of the ED pathway had no effect on the CAT productivity compared to the absence of knockouts (Fig.~\ref{fig:norm}A).
Pairwise knockouts of the ED pathway and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) also resulted in the same optimal CAT productivity.
However, there was a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); thus, alternative optimal metabolic flux distributions exist for CAT production, despite experimental constraints.
In addition, knockouts of amino acid biosynthesis reactions had no effect on the productivity with the exception of ALA, ASP, ASN and GLU, GLN biosynthesis reactions, since amino acids were supplemented in the media.
ssFBA rerouted the metabolic flux to have the same productivity but an alterantive flux distribution (alternative optima), as can be seen for a pairwise knockout of GLY, SER biosynthesis with HIS biosynthesis (Fig.~\ref{fig:norm}B).
Ultimately, to determine the metabolic flux distribution occurring in CFPS, we need to add additional constraints to the flux estimation calculation.
For example, thermodynamic feasibility constraints may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013},
and $^{13}$C labeling in CFPS could provide significant insight.
However, while $^{13}$C labeling techniques are well established for \emph{in~vivo} processes \cite{Zamboni:2009},
application of these techniques to CFPS remains an active area of research.

@@

\subsection*{Formulation and solution of the model equations.}
The sequence-specific flux balance analysis problem was formulated as a linear program:
\begin{equation}
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{X} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\mathcal{L}_{i} \leq w_i \leq \mathcal{U}_{i}  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective cost vector
and $\mathcal{L}_{i}$ and $\mathcal{U}_{i}$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The transcription (T) and translation (X) stoichiometry was modeled based upon the template reactions of Allen and Palsson \cite{Allen:2003aa} (Table \ref{tbl:template-reactions}).
The objective of the sequence specific flux balance calculation was to maximize the rate of protein translation, $w_{X}$.
The total glucose uptake rate was bounded by [0,40 mM/h] according to experimental data, while the amino acid uptake rates were bounded by [0,30 mM/h], but did not reach the maximum flux.
Gene and protein sequences were taken from literature and are available in the Supporting Information.
The sequence specific flux balance linear program was solved using the GNU Linear Programming Kit (GLPK) v4.55 \cite{GLPK}.
For all cases, amino acid degradation reactions were blocked as these enzymes were knocked out during the cell-free extract preparation \cite{2005_calhoun_BiotechnologyProgress, Garamella:2016aa}.
In the second case, all amino acid synthesis reactions were set to 0 mM/hr since \textit{E. coli} was grown in the presence of amino acids, thus these enzymes would not be present in the cell-free extract media.
In the third case, amino acid uptake reactions were set to 0 mM/hr.
In the experimental constrained case, \textit{E. coli} was grown in the presence of 13 amino acids (alanine, arginine, cysteine, serine, aspartate, glutamate, and glutamine were excluded) \cite{Zawada:2003}, thus the synthesis reactions responsible for those 13 amino acid were set to 0 mM/hr.

@@

\subsection*{Calculation of energy efficiency.}
Energy efficiency ($\mathcal{E}$) was calculated as the ratio of protein production to glucose consumption, written in terms of equivalent ATP molecules:
\begin{equation}\label{eqn:energy-efficiency-definition}
	\mathcal{E}\equiv\frac{w_{X}\left(2\cdot\rm {ATP}_{X}+\rm {GTP}_{X}\right)+w_{T}\left(2\times\left(\rm {ATP}_{T}+\rm {CTP}_{T}+\rm {GTP}_{T}+\rm {UTP}_{T}\right) \right)}{\texttt{q}_{GLC}\cdot\texttt{ATP}_{GLC}}
\end{equation}
where $\rm {ATP}_{T}$, $\rm {CTP}_{T}$, $\rm {GTP}_{T}$, $\rm {UTP}_{T}$ denote the stoichiometric coefficients of each energy species for the transcription of the protein of interest, $\rm {ATP}_{X}$, $\rm {GTP}_{X}$ denote the stoichiometric coefficients of ATP and GTP for the translation of the protein of interest, $\texttt{q}_{GLC}=w_{GLC}$ denotes the glucose uptake rate,
and $\texttt{ATP}_{GLC}$ denotes the equivalent ATP number for glucose produced.
The energy species stoichiometric coefficients are available in the Supporting Information.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
